---
title: "Hotel Bookings Analysis: Families with & without Children"
subtitle: "An EDA for Targeted Marketing Spend 2022"
author: "Jose Oviedo"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(knitr)
```


```{r, include=FALSE}
##load dataset from .csv file

hotel_bookings <- read_csv("hotel_bookings.csv")
```

# Intro to Market Segments

This geom_bar() plot shows how much each market segment contributes to overall hotel stays. From here we can tell that Online Travel Agency bookings are a significant market segment making up hotel bookings followed by Offline Travel Agency and Offline Travel Organizations (such as Costco Travel) and Group travel(I assume youth sports teams, etc.).

```{r}
##geom_bar() viz with fill aes to differentiate a variable, then faceted into individual plots

ggplot(hotel_bookings) +
  geom_bar(mapping = aes(x=hotel, fill=market_segment)) +
  facet_wrap(~market_segment)
```



```{r, include=FALSE}
##filter resort hotels/Online TA

onlineLT_resort_hotel<- hotel_bookings |>
  filter(hotel== "Resort Hotel") |>
  filter(market_segment== "Online TA")
```

## Resort Hotel Analysis

We then filter our dataset and isolate Resort Hotels that are Booked Online through a Travel Agency as this is the largest market segment to target with lead time focused marketing spend. In this plot we see that families with no children and 1 or 2 children are most likely to book online with a travel agency with relatively the same lead time. However, families with 3 or more children book at the last minute as a whole. So for targeted marketing spend, the data shows it is best to target larger families with children at the last minute before expected bookings.

```{r}
##plot Resort Hotel/Online TA families lead time with & without children

ggplot(onlineLT_resort_hotel) +
  geom_point(mapping = aes(x=children, y= lead_time, color= children)) +
    labs(title = "Resort Hotels Onine TA/LT of Families w/ & w/o Children")
```


```{r, include=FALSE}
##filter City Hotel/Online TA

onlineLT_city_hotel<- hotel_bookings |>
  filter(hotel== "City Hotel") |>
  filter(market_segment== "Online TA")
```

## City Hotel Analysis

We then filter our dataset to isolate City Hotels and that are Booked Online through a Travel Agency as this is the largest market segment to target with lead time focused marketing spend. Just like Resort Hotel bookings we can see that families with no children and 1 or 2 children are most likely to book online with a travel agency with relatively the same lead time and families with 3 or more children book similarly at the last minute. So for targeted marketing spend, the data shows it is best to target larger families with children at the last minute before expected bookings.

```{r}
##plot City Hotel/Online TA families lead time with & without children

ggplot(onlineLT_city_hotel) +
  geom_point(mapping = aes(x=children, y= lead_time, color= children)) +
    labs(title = "City Hotels Online TA/LT of Families w/ & w/o Children")
```


## Conclusion for Targeted Marketing Spend Q4 2022

For both Resort Hotels and City Hotels the data suggests it is best to spend marketing dollars on families with 3 or more children at the last minute before booking and for families with 2 or less children at any time with the same expected effectiveness.

