---
title: "The Ask: Does a higher GDP per capita mean a longer life expectancy? If so, what are the thresholds?"
subtitle: "Global Life Expectancy Trends Relative to Annual GDP per capita (50K or less and separated by Continent)"
author: "Data Vizzes by Jose Oviedo using R Programming (tidyverse, ggplot, gapminder, and other R packages)"
date: "February 2023"
output:
  prettydoc::html_pretty:
    theme: hpstr
    highlight: github
---

## Exploratory vizzes with gapminder data set:
*using tidyverse, gapminder, and other R packages*

```{r setup, message=FALSE, warning=FALSE, results='hide'}

library(tidyverse)
library(gapminder)
library(knitr)
library(rmarkdown)
library(prettydoc)
```

## Load gapminder data set and view to get a feel for it:
*using tidyverse, gapminder, and other R packages*

```{r explore-gapminder, message=FALSE, warning=FALSE, results='hide'}

data(gapminder)
view(gapminder)
attach(gapminder)
```

## Explore the data set:
*using tidyverse, gapminder, and other R packages*

```{r explore-gapminder2, message=FALSE, warning=FALSE, results='hide'}
## data set was collected from years 1952 to 2007

summary(gapminder)
```

```{r explore-gapminder3, message=FALSE, warning=FALSE}

glimpse(gapminder)
```

### Create Vizzes:
*using tidyverse, gapminder, and other R packages*

First, we will filter the *gapmminder* data set using our parameters for the 'gdpPercap' and 'year' variables.

Second, we will use these ggplot functions to make the graphic less busy and less messy.

  * geom_point(alpha= 0.3), transparancy of data points to improve readability.
  * geom_smooth(method = lm), linear modeling used 1952 and 2007 plots, but for plot that 'year' is aggregated I will not use linear modeling.
  * facet_wrap(), disaggregates the subset of 'continent' variable and shows each continent individually plotted.
  
```{r viz-1952, message=FALSE, warning=FALSE}
## log() to 'zoom' in on data points that are right-skewed
## plots do not need to be assigned to a df, will be exported or coded into Rmd file

## 1952 only

gapminder |> 
  filter(gdpPercap < 12000, year == 1952) |>
  ggplot(mapping = aes(gdpPercap, lifeExp, size = pop, color = continent)) +
    geom_point(alpha = 0.3) +
    geom_smooth(method = lm) +
    facet_wrap(~continent) + 
    labs(title = "1952 Global Life Expectancy Trends VS. Annual GDP per capita 12K or less separated by Continent",
       subtitle = ("Size of data points represents country population size"),
       caption = "Source: gapminder::gapminder data set from year 1952")
```

```{r viz-2007, message=FALSE, warning=FALSE}
## 2007 only

gapminder |> 
  filter(gdpPercap < 50000, year == 2007) |>
  ggplot(mapping = aes(gdpPercap, lifeExp, size = pop, color = continent)) +
  geom_point(alpha = 0.3) +
  geom_smooth(method = lm) +
  facet_wrap(~continent) + 
  labs(title = "2007 Global Life Expectancy Trends VS. Annual GDP per capita 50K or less separated by Continent",
       subtitle = ("Size of data points represents country population size"),
       caption = "Source: gapminder::gapminder data set from year 2007")
```

```{r viz-1952-to-2007, message=FALSE, warning=FALSE}
## 1952 to 2007 aggregated

gapminder |> 
  filter(gdpPercap < 50000, year > 1950) |>
  ggplot(mapping = aes(log(gdpPercap), lifeExp, size = pop, color = continent)) +
  geom_point(alpha = 0.3) +
  geom_smooth() +
  facet_wrap(~continent) + 
  labs(title = "1952-2007 Global Life Expectancy Trends VS. Annual GDP per capita 50K or less separated by Continent",
       subtitle = ("Size of data points represents country population size"),
       caption = "Source: gapminder::gapminder data set from years 1952 to 2007")
```

## Use ttest for the null hypothesis:
*using tidyverse, gapminder, and other R packages*

```{r summary-lm, message=FALSE, warning=FALSE, results='show'}
## lifeExp is the response/dependent variable
## gdpPercap is the explanatory variable 
## what we think might explain the changes in response variable

summary(lm(lifeExp ~ gdpPercap))
```


## Conclusions:
*using tidyverse, gapminder, and other R packages*

  *